LaMachine: A meta-distribution for NLP software

### 2019-01-17 docs/lamachine_clarin2018.tex
paper update

\section{Introduction} \label{intro}

@@

We already addressed the need for different interfaces for different audiences in Section~\ref{intro}. The challenge we
face, and for which LaMachine offers a solution, is one of software maintainability and distribution. How do we maintain
and distribute software that is often highly complex and consists of multiple interconnected components, considering
that this software is used differently by different audiences? A key aspect here is the reusability and accessibility of individual
software components. The philosophy we subscribe to encourages the development and distribution of software in a layered or modular fashion,
allowing each building block to serve as the foundation for another more high-level interface, without sacrificing the usability of the
foundation as such. This is in line with the UNIX philosophy of developing tools that do \emph{"one thing only, and do
it well"} and is contrasted with monolithic software solutions that are limiting because they either provide only high-level interfaces but do not
expose their inner components to build upon, or they are as a swiss-army knife full of needless but inseparable
components.

@@

At the start of the development cycle for LaMachine v2, we investigated whether Anaconda and its ecosystem would provide
a sufficient foundation for us to build upon. We decided this was not the case for a number of reasons: Anaconda
introduces more overhead than we desired for our purposes and conflicted with certain technology choices we made. We
wanted our native flavours (local environment and global environment) to be as close to the underlying distribution as possible and to reuse existing
technologies and packages from the distribution (or from homebrew for macOS), such as compilers (e.g. gcc) and
interpreters (e.g. python).  Anaconda, on the other hand, choses to provide its own packages and builds on those. It
does so using its own package manager and package repositories. This would require repackaging certain software for the
anaconda ecosystem, especially handling of non-Python software was not trivial.  Whereas LaMachine consists of a series
of Ansible scripts for the installation and configuration of software, Anaconda has its own package manager (conda)
that draw from its own package repositories (conda-forge).

@@

\section{Case study}\label{sec-case}

@@

@@

The recent release of LaMachine v2, which constituted a full rewrite, has opened up LaMachine to outside
contribution. Contributor documentation has been written, and at this stage, we greatly welcome external participants
to join in. Use cases as the example in section \ref{sec-case} contribute to thorough testing and running of LaMachine in less ideal circumstances such as nested VM constructions and offline usage.

@@

>>>>

\section{Introduction}
\label{sec:intro}

@@

We already addressed the need for different interfaces for different audiences in Section~\ref{sec:intro}. The challenge we
face, and for which LaMachine offers a solution, is one of software maintainability and distribution. How do we maintain
and distribute software that is often highly complex and consists of multiple interconnected components, considering
that this software is used differently by different audiences? A key aspect here is the reusability and accessibility of individual
software components. The philosophy we subscribe to encourages the development and distribution of software in a layered or modular fashion,
allowing each building block to serve as the foundation for another more high-level interface, without sacrificing the usability of the
foundation as such. This is in line with the UNIX philosophy of developing tools that do \emph{"one thing only, and do
it well"} and is contrasted with monolithic software solutions that are limiting because they either provide only high-level interfaces but do not
expose their inner components to build upon, or they are as a swiss-army knife full of needless but inseparable
components.

@@

At the start of the development cycle for LaMachine v2, we investigated whether Anaconda and its ecosystem would provide
a sufficient foundation for us to build upon. We concluded this was not the case for a number of reasons: Anaconda
introduces more overhead than we desired for our purposes and conflicted with certain technology choices we made. We
wanted our native flavours (local environment and global environment) to be as close to the underlying distribution as possible and to reuse existing
technologies and packages from the distribution (or from homebrew for macOS), such as compilers (e.g. gcc) and
interpreters (e.g. python).  Anaconda, on the other hand, choses to provide its own packages and builds on those. It
does so using its own package manager (conda) and package repositories (conda-forge). This would require repackaging certain software for the
anaconda ecosystem, especially handling of non-Python software was not trivial, as Anaconda was originally conveived as a
Python distribution.

@@

\section{Case studies}
\label{sec:case}

\subsection{Text Mining for Health Inspection}

@@

\subsection{Workshop: Cataloguing of Textual Cultural Heritage Objects}

%TODO

@@

LaMachine provides a flexible solution for the installation of a variety of NLP software, resulting in a kind
of virtual laboratory that, through various interfaces, can be employed by a variety of people, from developers and data
scientists to the wider research community that CLARIN explicitly addresses.

Use cases as the examples in section \ref{sec:case} contribute to thorough testing and running of LaMachine in less ideal
circumstances such as nested VM constructions and offline usage.

@@

LaMachine v2 is open to  outside contribution. Contributor documentation has been written, and at this stage, we greatly
welcome external participants to join in.
